---
title: "Accurate estimation of the prevalence of asymptomatic SARS-CoV-2 infections from population-based surveys"
---

```{r setup, include=FALSE}
rm(list = ls())
gc()

suppressPackageStartupMessages({
  library(tidyverse) 
  library(gridExtra)  # Arrange multiple ggplots
  library(reshape2)   
  library(ggrepel)    # Prevent text overlap in plots
})
```

# 1. Loading data
```{r, warning=FALSE}
setwd("/path/to/be/specified")
# Change this path to the directory where a data file will be stored on your system
df = read.csv('./data.csv', header = TRUE)  
# The data file can be downbloaded from this repository

```

Note: The variables in the following calculations are defined as in the main text.

# 2. Exploring data of serosurvey
## 2.1. Violin plots of rho_c (Figure S1) and phi_c (Figure 1a)

```{r}
# Define styling parameters
s_border <- 0.75
s_point <- 1.5

# Convert data to long format for violin plot
df_long <- melt(df, measure.vars = c("rho_c", "phi_c"))

# Function to create violin plots
plot_violin <- function(data, var_name, fill_color) {
  ggplot(data %>% filter(variable == var_name), aes(x = variable, y = value)) +
    geom_violin(trim = TRUE, scale = "width",
                width = 1, size = 1, alpha = 0.3, fill = fill_color) +
    geom_boxplot(width = 0.5, color = "black", outlier.shape = NA, 
                 alpha = 0.5, fill = fill_color, size = 1, fatten = 1.25) +
    geom_jitter(shape = 21, width = 0.2, size = s_point, fill = "gray",
                color = "black", alpha = 0.6) +
    theme_minimal() +
    theme(axis.title = element_blank(),
          axis.text = element_text(size = 11, colour = "black"),
          axis.ticks = element_line(linewidth = 0.5, colour = "black"),
          panel.border = element_rect(colour = "black", fill = NA, size = s_border),
          legend.position = "none")
}

# Plot violin for rho_c
plot_violin(df_long, "rho_c", "olivedrab3")

# Plot violin for phi_c
plot_violin(df_long, "phi_c", "orchid3") +
  theme(plot.margin = margin(5, 10, 5, 5))

# Print medians of rho_c and phi_c
cat("Median of rho_c:", median(df$rho_c), "\n")
cat("Median of phi_c:", median(df$phi_c), "\n")

```
# 3. Estimates of the prevalence of asymptomatic infections
## 3.1. Computing corrected prevalence, psi, and percentage increase in the prevalence, eta

```{r}
# Compute psi using Equation (1) from the main text
df$psi <- with(df, {
  rho <- (rho_c + beta - 1)/(alpha + beta - 1)
  numerator <- rho_c*(1-rho_c)*(psi_c - 1 + phi_c)
  denominator <- rho*(psi_c*rho_c*(1-alpha) - alpha*(1-phi_c)*(1-rho_c))
  1 - numerator/denominator
})

# Compute Wilson confidence intervals for ψ (used in Figure 1b)
a <- df$psi + ((1.96^2) / (2 * df$sp))  # sp: number of seropositive cases
b <- 1.96 * sqrt(((df$psi * (1 - df$psi)) + (1.96^2 / (4 * df$sp))) / df$sp)
c <- 1 + (1.96^2 / df$sp)

C_l <- (a - b)/c
C_u <- (a + b)/c

df$ci_lo <- C_l
df$ci_up <- C_u

# eta
df$eta <- 100*(df$psi - df$psi_c)/df$psi_c

```

## 3.2. Comparing psi and psi_c

```{r}
# Count number of estimates where psi >= psi_c
print(length(which(df$psi >= df$psi_c)))

# Wilcoxon signed-rank test for paired comparison of psi and psi_c
wilcox_test <- wilcox.test(df$psi - df$psi_c, 
                           mu = 0, 
                           alternative = "greater", 
                           na.rm = TRUE, 
                           conf.int = 0.95)
print(wilcox_test)

# Compute median values of psi_c and psi
print(median(df$psi_c, na.rm = TRUE))
print(median(df$psi, na.rm = TRUE))

```

### 3.2.1. Scatter plot of psi versus psi_c (Figure 1b)

```{r}
# Compute quantiles of η (used for color gradient in Figure 1b)
quan_eta <- quantile(df$eta)

# Create plot
plt <- ggplot(df) +
  geom_line(aes(x = psi_c, y = psi_c), color = "black", linetype = "dashed") +
  geom_errorbar(aes(x = psi_c, ymin = C_l, ymax = C_u),   
                color = "black", width = 0.01, size = 0.3, alpha = 0.8) +
  geom_point(aes(x = psi_c, y = psi, fill = eta), 
             shape = 21, color = "black", size = s_point) +
  # Annotate country labels using geom_text_repel()
  geom_text_repel(data = df, aes(x = psi_c, y = psi, label = country, 
                                  colour = eta), angle = 90, max.overlaps = 15) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.25), limits = c(0, 1.02), 
                     expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25), limits = c(0, 1.03), 
                     expand = c(0, 0)) +
  scale_fill_gradientn(colours = c("darkviolet", "darkblue", "dodgerblue", 
                                   "darkgreen", "gold", "chocolate", "firebrick3"),
                       values = scales::rescale(c(-80, 6, 30, 76, 377))) +  
  # The breaks come from the quantiles of eta
  
  scale_colour_gradientn(colours = c("darkviolet", "darkblue", "dodgerblue", 
                                     "darkgreen", "gold", "chocolate", "firebrick3"),
                         values = scales::rescale(c(-80, 6, 30, 76, 377))) + 
  labs(x = "psi_c", y = "psi") +
  theme_minimal() +
  theme(axis.text = element_text(size = 11, colour = "black"),
        axis.ticks = element_line(linewidth = 0.5, colour = "black"),
        panel.border = element_rect(colour = "black", fill = NA, size = s_border),
        legend.position = "none")

plt

```

### 3.2.2. Violin plots of psi_c and psi (Figure 1c)

```{r}
# Convert data to long format for violin plot
df_long <- melt(df, measure.vars = c("psi_c", "psi"))

# Generate violin plot
plt <- ggplot(df_long[df_long$variable %in% c("psi_c", "psi"), ],
             aes(x = variable, y = value, fill = variable)) +
  geom_violin(position = position_dodge(width = 0.8), 
              trim = TRUE, scale = "width",
              width = 0.8, size = 1, alpha = 0.3) +
  geom_boxplot(width = 0.4, color = "black", outlier.shape = NA, 
               alpha = 0.5, size = 1, fatten=1.25) +
  scale_fill_manual("values" = c("psi_c" = "orange", "psi" = "red")) +
  geom_jitter(shape = 21, width = 0.2, size = s_point, color = "black") +
  theme_minimal() +
  theme(axis.text = element_text(size = 11, colour = "black"),
        axis.title = element_blank(),
        axis.ticks = element_line(linewidth = 0.5, colour = "black"),
        panel.border = element_rect(colour = "black", fill = NA, 
                                    size = s_border),
        plot.margin = margin(5, 15, 5, 15),
        legend.position = "none")
plt
```

### 3.2.3. Histogram of eta (Figure 1d)

```{r}
# Define categorical bins for eta
df$eta_range <- cut(df$eta,
                    breaks = c(-Inf, 0, 25, 50, 75, 100, Inf),
                    labels = c("0", "0-25", "25-50", "50-75", "75-100", "> 100"),
                    right = FALSE)  # Intervals include left endpoint and exclude right

# Create histogram
plt <- ggplot(df, aes(x = eta_range)) +
  geom_bar(fill = "skyblue", color = "black", width = 0.8) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 11, colour = "black", angle = 45,
                                   hjust = 1),
        axis.text.y = element_text(size = 11, colour = "black"),
        axis.ticks = element_line(linewidth = 0.5, colour = "black"),
        panel.border = element_rect(colour = "black", fill = NA, 
                                    size =s_border),
        plot.margin = margin(5, 5, 5, 10),
        legend.position = "none")

plt
```
# 4. Factors contributing to the correction in the estimate of prevalence
## 4.1. Correlation between predictors of eta (phi_c, beta, alpha, and rho_c) (Figure S2)

```{r, warning=FALSE}
# Scatter plots

s_point = 1.5
s_font = 11

## alpha and phi_c
plt1 <- ggplot(df, aes(x = phi_c, y = alpha)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text.y = element_text(size = s_font, colour = "black"),
        axis.ticks.y = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA,
                                    size=s_border),
        legend.position = "none")

plt1

## beta and phi_c
plt2 <- ggplot(df, aes(x = phi_c, y = beta)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text.y = element_text(size = s_font, colour = "black"),
        axis.ticks.y = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA,
                                    size=s_border),
        legend.position = "none")

plt2

## rho_c and phi_c
plt3 <- ggplot(df, aes(x = phi_c, y = rho_c)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text = element_text(size = s_font, colour = "black"),
        axis.ticks = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA,
                                    size=s_border),
        legend.position = "none")

plt3

## beta and alpha
plt4 <- ggplot(df, aes(x = alpha, y = beta)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(panel.border = element_rect(colour = "black", fill=NA,
                                    size=s_border),
        legend.position = "none")

plt4

## rho_c and alpha
plt5 <- ggplot(df, aes(x = alpha, y = rho_c)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text.x = element_text(size = s_font, colour = "black"),
        axis.ticks.x = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA,
                                    size=s_border),
        legend.position = "none")

plt5

## rho_c and beta
plt6 <- ggplot(df, aes(x = beta, y = rho_c)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text.x = element_text(size = s_font, colour = "black"),
        axis.ticks.x = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA,
                                    size=s_border),
        legend.position = "none")

plt6

# Compute Spearman's correlation coefficients
attach(df)
cor1 <- cor.test(alpha, phi_c, method = "spearman")
cor1
cor2 <- cor.test(beta, phi_c, method = "spearman")
cor2
cor3 <- cor.test(rho_c, phi_c, method = "spearman")
cor3
cor4 <- cor.test(beta, alpha, method = "spearman")
cor4
cor5 <- cor.test(rho_c, alpha, method = "spearman")
cor5
cor6 <- cor.test(rho_c, beta, method = "spearman")
cor6

```
## 4.2. Multiple linear regression of eta on predictors
```{r}
model <- lm(eta ~ phi_c + beta + alpha + rho_c, data = df)
summary(model)
summary(model)$r.squared         # R-squared value
summary(model)$coefficients[,4]    # p-values of predictors

```
## 4.3. Correlation between eta and predictors (Figure S3 and Figure 2)
```{r}
# Compute correlation coefficients
attach(df)
cor1 <- cor.test(beta, eta, method = "spearman")
cor1
cor2 <- cor.test(alpha, eta, method = "spearman")
cor2
cor3 <- cor.test(rho_c, eta, method = "spearman")
cor3
cor4 <- cor.test(phi_c, eta, method = "spearman")   # Used in Figure 2
cor4
cor4$p.value    

# Scatter plots
s_point = 1.5
s_font = 11

## eta and beta
plt1 <- ggplot(df, aes(x = beta, y = eta)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text = element_text(size = s_font, colour = "black"),
        axis.ticks = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA, size=s_border),
        legend.position = "none")

plt1

## eta and alpha
plt2 <- ggplot(df, aes(x = alpha, y = eta)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text = element_text(size = s_font, colour = "black"),
        axis.ticks = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA, size=s_border),
        legend.position = "none")

plt2

## eta and rho_c
plt3 <- ggplot(df, aes(x = rho_c, y = eta)) +
  geom_point(size = s_point, color= "black") +
  theme_bw() +
  theme(axis.text = element_text(size = s_font, colour = "black"),
        axis.ticks = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA, size=s_border),
        legend.position = "none")

plt3

```
## 4.4. Dependence of eta on phi_c

```{r}
# Model showing dependence of eta on phi_c for a perfect test (Equation (2) of the main text), alpha = beta = 1
df$eta_model <- 100*(df$phi_c/(1-df$phi_c))

# Model-based prediction on data of all the serosurveys included in our work
## Compute residual sum of squares (RSS) and total sum of squares (TSS)
rss <- sum((df$eta_model - df$eta)^2)
tss <- sum((df$eta - mean(df$eta))^2)

## Compute R-squared value
r_sq <- 1-rss/tss
r_sq
```

## 4.5. Model Prediction of Serosurvey Data (Figure 2)

```{r}
# Generate model-based prediction
phi_c_model = seq(min(df$phi_c), max(df$phi_c), length.out = 1000)

df_model <- data.frame(phi_c_model = phi_c_model,
                       eta_model = 100*phi_c_model/(1-phi_c_model))
# eta_model: eta for a perfect test, obtained from equation (2) of the main text

# Create plot
plt <- ggplot() +
  geom_line(data = df_model, aes(x = phi_c_model, y = eta_model), color = "black",
            linewidth = 1) +   # eta_model
  geom_point(data = df, aes(x = phi_c, y = eta, fill = eta), shape=21, color = "black",
             size = s_point, position = "jitter") +   # eta from data of quantities in Q from all the serosurveys
  geom_text_repel(data = df, aes(x = phi_c, y = eta, label = country, color = eta),
                  angle = 90,
                  nudge_x = 0, 
                  size = 4,
                  min.segment.length = 1,
                  max.overlaps = Inf) +
  scale_fill_gradientn(colours = c("darkviolet", "darkblue", "dodgerblue", 
                                   "darkgreen", "gold", "chocolate", "firebrick3"),
                       values = scales::rescale(c(-80, 6, 30, 76, 377))) +
  scale_colour_gradientn(colours = c("darkviolet", "darkblue", "dodgerblue", 
                                     "darkgreen", "gold", "chocolate", "firebrick3"),
                         values = scales::rescale(c(-80, 6, 30, 76, 377))) + 
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(size = 11, colour = "black"),
        axis.ticks = element_line(linewidth = 0.5, colour="black"),
        panel.border = element_rect(colour = "black", fill=NA, size=s_border),
        legend.position = "none")

plt
```

# 5. Improved consistency between estimates of prevalence
## 5.1. Dataset of countries with multiple studies

```{r}
# Extracting countries where multiple studies were available
df_multiple <- df %>% 
  group_by(country) %>% 
  filter(n() > 1) %>% 
  ungroup()

# Obtaining variance of psi_c and psi within countries with multiple studies
country_variance <- df_multiple %>% 
  group_by(country) %>% 
  summarise(var_c = var(psi_c),
            var = var(psi),
            var_red = (var(psi_c)-var(psi))*100/var(psi_c))  
## var_red: Percentage reduction in variance before and after correction
# view(country_variance)

# Perform a one-tailed paired t-test to test if variance reduction is significant
t_test_result <- t.test(country_variance$var, country_variance$var_c, 
                        paired = TRUE, alternative = "less")
t_test_result

```
## 5.2. Dataset of all countries (with single or multiple studies)

```{r}
library(car)

# Percentage reduction in variance following correction
print((var(df$psi_c) - var(df$psi))/var(df$psi_c))

# Combine data into a single dataframe to do significance test
df1 <- data.frame(
  value = c(df$psi_c, df$psi),
  group = rep(c("psi_c", "psi"), times = c(length(df$psi_c), length(df$psi)))
)

# Perform Levene's test
levene_test_result <- leveneTest(value ~ group, data = df1)

# Display results
print(levene_test_result)

```
